---
title: "EDA"
output: html_document
---
# Load data
```{r}
library(tidyverse)
library(tidytuesdayR)
```


```{r}
tuesdata <- tidytuesdayR::tt_load(2025, week = 23)

judges_appointments <- tuesdata$judges_appointments |>
  mutate(across(ends_with("date"), mdy))
judges_people <- tuesdata$judges_people
```

```{r}
#write_csv(judges_appointments, "judges_appointments.csv")
```


# Look at each column

```{r}
judges_appointments |>
  select(judge_id) |>
  print(n = 100)


complete_sequence <- tibble(judge_id = seq(from = min(judges_appointments$judge_id), to = max(judges_appointments$judge_id)))

complete_sequence |>
  anti_join(judges_appointments) |>
  anti_join(complete_sequence |> anti_join(judges_people))

complete_sequence |>
  anti_join(judges_people) |>
  print(n = 50)

judges_appointments |>
  count(judge_id, sort = T) |>
  filter(n > 1) |>
  print(n = Inf)

judges_appointments |>
  filter(judge_id == 2689) |>
  View()

judges_appointments |>
  count(court_name, sort = TRUE) |>
  View()

judges_appointments |>
  count(court_type, sort = TRUE)

judges_appointments |>
  count(president_name, sort = TRUE) |>
  print(n = 50)

length(unique(judges_appointments$president_name))

judges_appointments |>
  filter(!president_party %in% c("Reassignment", "Assignment")) |>
  distinct(president_name, president_party) |>
  arrange(president_name) |>
  print(n = 50)
```

# How many judges have Pacific Islander as part of their designated race?

```{r}
judges_people |>
  count(race)
```

# Which Presidents appointed the most judges? The fewest?

```{r}
judges_appointments |>
  group_by(president_name) |>
  summarise(
    unique_judges = n_distinct(judge_id),
    n = n()
  ) |>
  arrange(desc(unique_judges))
```

# Which political parties have appointed the most judges to courts of customs or internation trade? Due to some coding issues in the current version of the {historydata} package, you may need to combine some terms to find all such appointments.

```{r}
judges_appointments |>
  filter(court_type %in% c("U. S. Court of Customs and Patent Appeals", "U. S. Court of International Trade", "U. S. Court of Inter", "U. S. Court of Custo")) |>
  count(president_party, sort = TRUE)
```

# Making some kind of plot

```{r}
judges_appointments |>
  filter(!president_name %in% c("Assignment", "Reassignment")) |>
  group_by(president_name, president_party) |>
  summarise(
    min_date = min(nomination_date, na.rm = T),
    max_date = max(nomination_date, na.rm = T),
    n = n()
  ) |>
  arrange(max_date) |>
  print(n = 50) |>
  ggplot(aes(reorder(president_name, max_date), n, fill = president_party)) +
  geom_col() +
  coord_flip()

judges_appointments |>
  ggplot(aes(4 * (year(nomination_date) %/% 4))) +
  geom_bar()

judges_appointments |>
  filter(president_name == "John F. Kennedy") |>
  View()

judges_appointments |>
  filter(!president_name %in% c("Assignment", "Reassignment")) |>
  group_by(president_name) |>
  mutate(max_date = max(nomination_date)) |>
  ungroup() |>
  mutate(president_num = dense_rank(desc(max_date))) |>
  filter(president_num <= 10) |>
  ggplot(aes(reorder(president_name, nomination_date, .na_rm = TRUE), fill = president_party)) +
  geom_bar() +
  coord_flip() +
  labs(
    x = "President name",
    y = "# judges appointed",
    fill = "Party",
    title = "Judge appointments last 10 presidential terms"
  )
```

```{r}
ggsave(
  filename = "judges-appointed-last-10.png",
  height = 5,
  width = 7,
  units = "in",
  dpi = 300
)
```







