---
title: "EDA"
output: html_document
---

```{r}
library(tidyverse)
library(tidytuesdayR)
```

```{r}
tuesdata <- tidytuesdayR::tt_load(2025, week = 20)

water_quality <- tuesdata$water_quality |>
  rename(lat = latitude, lon = longitude)
weather <- tuesdata$weather |>
  rename(lat = latitude, lon = longitude)
```

```{r}
water_quality |>
  group_by(region) |>
  summarise(max_1 = max(lat), min_1 = min(lat), max_2 = max(lon), min_2 = min(lon))

water_quality |>
  distinct(region, swim_site) |>
  arrange(swim_site)

water_quality |>
  distinct(council, swim_site) |>
  group_by(swim_site) |>
  filter(n() == 1)

water_quality |>
  distinct(swim_site, lat, lon) |>
  group_by(swim_site) |>
  filter(n() > 1)

water_quality |>
  group_by(swim_site) |>
  summarise(
    min_date = min(date),
    max_date = max(date)
  )

water_quality |>
  filter(swim_site == "Camp Cove")

summary(water_quality$enterococci_cfu_100ml)

water_quality |>
  filter(enterococci_cfu_100ml %% 1 != 0)

summary(water_quality$water_temperature_c)

water_quality |>
  filter(water_temperature_c < 9.5 | water_temperature_c > 29.5) |>
  print(n = Inf)
  
summary(water_quality$conductivity_ms_cm)

water_quality |>
  slice_sample(n = 10000) |>
  ggplot(aes(conductivity_ms_cm)) +
  geom_histogram()
``` 

```{r}
weather |>
  select(precipitation_mm) |>
  filter(precipitation_mm > 30)
  print(n = 50)
```

# Has the water quality declined over this period
```{r}
little_manly_cove <- water_quality |>
  filter(swim_site == "Little Manly Cove")
```

```{r}
library(lubridate)

summary(little_manly_cove$date)

little_manly_cove |>
  mutate(year = year(date)) |>
  group_by(year) |>
  summarise(
    avg = mean(enterococci_cfu_100ml),
    n = n()
  ) |>
  ggplot(aes(year, avg)) +
    geom_point() +
  scale_y_log10() +
  geom_smooth()

water_quality |>
  mutate(year = year(date)) |>
  group_by(year) |>
  summarise(
    avg = mean(enterococci_cfu_100ml, na.rm = TRUE),
    n = n()
  ) |>
  ggplot(aes(year, avg)) +
    geom_point() +
  scale_y_log10() +
  geom_smooth() +
  labs(
    subtitle = "Enterococci level declined over the years but is starting to increase again",
    x = "Year",
    y = "Enterococci level"
  )

ggsave(
  filename = "bacteria-level.png",
  height = 4.5,
  width = 6,
  units = "in",
  dpi = 300
)
```

# How does rainfall impact E-coli bacteria levels?

```{r}
average_water_quality_2025 <- water_quality |>
  filter(year(date) == 2025, !is.na(enterococci_cfu_100ml)) |>
  group_by(date) |>
  summarise(
    avg_enterococci_cfu = mean(enterococci_cfu_100ml),
    n = n()
  )

average_water_quality_2025 |>
  left_join(weather |> select(date, precipitation_mm), by = c("date")) |>
  pivot_longer(cols = c(avg_enterococci_cfu, precipitation_mm), 
               names_to = "metric", 
               values_to = "value") |>
  ggplot(aes(date, value, color = metric)) +
  geom_line() +
  scale_x_date(breaks = "15 days", date_labels = "%m/%d") +
  scale_y_log10() +
  labs(
    x = "Date",
    y = "",
    tittle = "Enterococci level seems to collerate with percipitation",
    subtitle = "Every spike in precipitation seems to predict a spike in enterococci level"
  ) 

ggsave(
  filename = "precipitation_vs_enterococci_level.png",
  height = 4.5,
  width = 6,
  units = "in",
  dpi = 300
)
```

```{r}
weather |>
  filter(year(date) == 2025) |>
  ggplot(aes(date, precipitation_mm)) +
  geom_line()

water_quality |>
  filter(year(date) == 2025, !is.na(enterococci_cfu_100ml)) |>
  left_join(weather |> select(date, precipitation_mm), by = c("date")) |>
  ggplot(aes(enterococci_cfu_100ml, precipitation_mm)) +
  geom_point()
```

# Are some swimming sites particularly prone to high bacteria levels following rain?

```{r}
average_water_quality_2025 |>
  left_join(weather |> select(date, precipitation_mm), by = c("date")) |>
  pivot_longer(cols = c(avg_enterococci_cfu, precipitation_mm), 
               names_to = "metric", 
               values_to = "value")

water_quality |>
  count(lat, lon, sort = TRUE)

water_quality |>
  filter(year(date) == 2025, !is.na(enterococci_cfu_100ml)) |>
  group_by(region, date) |>
  summarise(
    avg_enterococci_cfu = mean(enterococci_cfu_100ml),
    n = n()
  ) |>
  left_join(weather |> select(date, precipitation_mm), by = c("date"), relationship = "many-to-one") |>
  pivot_longer(cols = c(avg_enterococci_cfu, precipitation_mm), 
               names_to = "metric", 
               values_to = "value") |>
  ggplot(aes(date, value, color = metric)) +
  geom_line() +
  facet_wrap(~ region) +
  scale_x_date(breaks = "15 days", date_labels = "%m/%d") +
  scale_y_log10()
```

# Analyze bacteria levels against rain in different regions

```{r}
water_quality |>
  group_by(region) |>
  summarise(
    swim_site_count = n_distinct(swim_site),
    n = n()
  )

water_quality |>
  filter(region == "Western Sydney", year(date) == 2025, !is.na(enterococci_cfu_100ml)) |>
  group_by(swim_site, date) |>
  summarise(
    avg_enterococci_cfu = mean(enterococci_cfu_100ml),
    n = n()
  ) |>
  left_join(weather |> select(date, precipitation_mm), by = c("date"), relationship = "many-to-one") |>
  pivot_longer(cols = c(avg_enterococci_cfu, precipitation_mm), 
               names_to = "metric", 
               values_to = "value") |>
  ggplot(aes(date, value, color = metric)) +
  geom_line() +
  facet_wrap(~ swim_site) +
  scale_x_date(breaks = "15 days", date_labels = "%m/%d") +
  scale_y_log10()
```

# Swim sites in Southern Sydney are particularly prone to high bateria levels following rain

```{r}
water_quality |>
  filter(region == "Southern Sydney", year(date) == 2025, !is.na(enterococci_cfu_100ml)) |>
  group_by(swim_site, date) |>
  summarise(
    avg_enterococci_cfu = mean(enterococci_cfu_100ml),
    n = n()
  ) |>
  left_join(weather |> select(date, precipitation_mm), by = c("date"), relationship = "many-to-one") |>
  pivot_longer(cols = c(avg_enterococci_cfu, precipitation_mm), 
               names_to = "metric", 
               values_to = "value") |>
  ggplot(aes(date, value, color = metric)) +
  geom_line() +
  facet_wrap(~ swim_site) +
  scale_x_date(breaks = "15 days", date_labels = "%m/%d") +
  scale_y_log10()
```

# How about swim sites in Sydney City?

```{r}
water_quality |>
  filter(region == "Sydney City", year(date) == 2025, !is.na(enterococci_cfu_100ml)) |>
  group_by(swim_site, date) |>
  summarise(
    avg_enterococci_cfu = mean(enterococci_cfu_100ml),
    n = n()
  ) |>
  left_join(weather |> select(date, precipitation_mm), by = c("date"), relationship = "many-to-one") |>
  pivot_longer(cols = c(avg_enterococci_cfu, precipitation_mm), 
               names_to = "metric", 
               values_to = "value") |>
  ggplot(aes(date, value, color = metric)) +
  geom_line() +
  facet_wrap(~ swim_site) +
  scale_x_date(breaks = "15 days", date_labels = "%m/%d") +
  scale_y_log10()
```


```{r}
water_quality |>
  filter(region == "Northern Sydney", year(date) == 2025, !is.na(enterococci_cfu_100ml)) |>
  group_by(swim_site, date) |>
  summarise(
    avg_enterococci_cfu = mean(enterococci_cfu_100ml),
    n = n()
  ) |>
  left_join(weather |> select(date, precipitation_mm), by = c("date"), relationship = "many-to-one") |>
  pivot_longer(cols = c(avg_enterococci_cfu, precipitation_mm), 
               names_to = "metric", 
               values_to = "value") |>
  ggplot(aes(date, value, color = metric)) +
  geom_line() +
  facet_wrap(~ swim_site) +
  scale_x_date(breaks = "15 days", date_labels = "%m/%d") +
  scale_y_log10()
```












